\section{Evaluation Metrics}
Metrics used to evaluate lens properties include:
\begin{itemize}
    \item \textbf{Sharpness}: The sharpness evaluation using MTF charts showed that the application could accurately measure the lens's resolving power.
    \item \textbf{Bokeh}: The bokeh analysis revealed that the application effectively assessed the quality of out-of-focus areas.
    \item \textbf{Distortions}: Evaluated using test charts to identify barrel and pincushion distortion.
    \item \textbf{Chromatic Aberration}: Chromatic aberration analysis showed that the application could detect chromatic aberrations based on the misalignment of RGB components.
    \item \textbf{Vignetting}: The vignetting measurement revealed that the application could accurately compare luminance differences across the image.
\end{itemize}

\section{Comparative Analysis}
Results obtained from the application were compared with those from Imatest and Dxomark. The application provided comparable accuracy in certain evaluating lens properties, demonstrating its potential as a cost-effective and accessible tool for photographers.

\section{Discussion}

The comparative analysis highlights the application's success in evaluating various lens properties. By providing results comparable to industry-standard tools like Imatest and Dxomark, the application proves to be a valuable resource for photographers seeking to assess lens quality without excessive costs.

The success of the application in accurately measuring sharpness, bokeh, chromatic aberration, and vignetting demonstrates its comprehensive capabilities. The use of advanced image processing techniques and detailed analysis ensures that photographers can rely on the application for precise and detailed lens evaluations.

Overall, the application offers a cost-effective and accessible alternative to professional lens testing tools, making it an attractive option for both amateur and professional photographers. Future developments could focus on enhancing user interface features, expanding compatibility with different camera models, and incorporating additional lens evaluation metrics.
